# Reproducible Research  
## Peer assessment 1  
By Pixel81  
Date: "Monday, July 13, 2015"  
================================================================================

This document is a reproducible research made in the Coursera class
'Reproducible research'.  
In order to be readable, echo is set on true for all the code.

### Loading and preprocessing the data

To well work, all the work is done in the R Markdown file folder. The zip file
need to be download in this folder to properly work.  
First, we unzip the data from the zip file and load it into a variable.
```{r loaddata,echo = TRUE}

if(!file.exists('activity.csv')){ 
        unzip('repdata-data-activity.zip') 
        } # Unzip file
activity <- read.csv('activity.csv') # Load data in variable
```

We make a summary to have a quick overview of the file content.
```{r summary,echo = TRUE}
summary(activity)
```

To correctly interpretate date, we convert it with the POSIXct function. We also
add a date.time variable with the interval of 5 minutes in the date stream.
```{r convertdate,echo = TRUE}
activity$date <- as.POSIXct(activity$date,
                            format='%Y-%m-%d',
                            tz='GMT')
time <- formatC(activity$interval / 100, 2, format='f')
activity$date.time <- as.POSIXct(paste(activity$date, time),
                                 format='%Y-%m-%d %H.%M',
                                 tz='GMT')
```

Just take a look at our new content of activity.
```{r head,echo = TRUE}
head(activity)
```

### What is mean total number of steps taken per day ?
1. Calculate the total number of steps taken per day
```{r stepsperday, echo = TRUE}
total.steps <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)
```

2. Make a histogram of the total number of steps taken each day
```{r}
library(ggplot2)
ggplot(total.steps, aes(x=total.steps)) 
+ geom_histogram(binwidth=30, colour="white")

```

```{r}
plot(unique(activity$date),
     total.steps,
     main = "Steps taken each day",
     xlab = "Date",
     ylab = "Frequency (steps)",
     type = "h",
     lwd = 8,
     col = "violet")
```

3. Calculate and report the mean and median of the total number of steps taken per day.  
The mean calculation is rounded to fit on report.
```{r}
mean.steps <- round(mean(total.steps),0)
median.steps <- median(total.steps)
```

The mean is `r mean.steps` and the median is `r median.steps`.

### What is the average daily activity pattern ?

```{r}
```

### Imputing missing values

```{r}
```

### Are there differences in activity patterns between weekdays and weekends ?

```{r}
```